---
title: "PCA"
author: "Ahmed Al Dhouani 129466"
date: "2025-11-29"
output: html_document
---

This code performs a Principal Component Analysis on standardized economic indicators (Share, Growth, logGDP) 
to reduce dimensionality and identify the main underlying patterns in the data.
It extracts and interprets the variance explained and loadings of each principal 
component, then visualizes countries(Oman, Suadi, UAE) on a two dimensional biplot to compare their economic structures and growth profiles.

USED DATA: "PCA_Final_Data.xlsx"

```{r}
library(readxl)
library(ggplot2)
library(factoextra)
library(dplyr)
```



```{r}
df <- read_excel("PCA_Final_Data.xlsx")
```



```{r}
# PCA only needs the variables without Country and Year.
pca_data <- df %>%
  select(Share, Growth, logGDP)

# In order to obtain accurate results, it is necessary to standardize because the variables have different units.
pca_scaled <- scale(pca_data)
```




```{r}
# Step 3: Run PCA
pca_model <- prcomp(pca_scaled, center = TRUE, scale. = TRUE)
summary(pca_model)


# This will give us:
#The percentage of variance explained by PC1
#The percentage of variance explained by PC2
#How much they explain together (often above 80%, which is excellent)
```



```{r}
# Step 4: Extracting the Loadings
pca_model$rotation

#This tells us:
#Does PC1 represent “share + output level”?
#Does PC2 represent “growth”?
#From this, we can deduce the meaning of each component.
```




```{r}
# Step 5: Add the country and year to the chart
df$Country <- factor(df$Country)
```



```{r}
# Step 6: Drawing the Biplot (Most Important Part)
# This drawing will place:
# Oman
# United Arab Emirates
# Saudi Arabia
# on a two-dimensional map.


fviz_pca_biplot(
  pca_model,
  geom.ind = "point",
  habillage = df$Country, 
  addEllipses = TRUE,    
  label = "var",        
  repel = TRUE
)
```


=

